---
title: "Advanced Sports Data Lesson 1: Aggregates review"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to take lots of little things and total them up into bigger things.
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(tidyverse)
library(glue)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.completion=FALSE)
```

# Review 

## The basics

Students taking SPMC 460 should already be familiar with the basics of data analysis with the Tidyverse.

Should. 

The problem with many of these students is they have the memories of squirrels on meth. Five minutes after they finish a class -- whoosh. It's right out of their heads.

So we begin with some basics to jog the memory. The first reminder is that all notebooks start with library imports. Just about every assignment in this class will use, at a minimum, `library(tidyverse)` and `library(tidymodels)`. 

We'll get to tidymodels later. For now, run this.

```{r load-tidyverse, exercise=TRUE}
library(tidyverse)
```
```{r load-tidyverse-solution}
library(tidyverse)
```
```{r load-tidyverse-check}
grade_this_code()
```

**From now on, if that's not the first line of your notebook, you're probably doing it wrong.**

## Loading data

The next thing we need to do is to get some data to work with. We do that by reading it in. In our case, we're going to read data from a csv file -- a comma-separated values file.

The CSV file we're going to read is every NBA player since the 1979-1980 season -- Larry Bird and Magic Johnson's rookie season.

You can download the data here if you want to use it in your own notebook. **For purposes of this exercise, you don't need to do this. The data is included here if you want to try this in your own notebook.**

```{r echo=FALSE, class.output="bg-info", results="asis",  message=FALSE,  warning=FALSE}
library(downloadthis)
library(glue)

dllink <- download_link(
  link = "http://mattwaite.github.io/sportsdatafiles/allplayers.csv",
  button_label = "Download csv file",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)

glue("<pre><p><strong>For this walkthrough:</strong></p><p>{dllink}</p></pre>")

```

A reminder of what you are going to do. First, you are going to give your data a name -- let's use `players`. Then, you'll use the arrow -- `<-` to assign the results of what comes next to the name `players`. What comes next? We'll use the `read_csv` function to ... read a csv file. What goes inside the () after `read_csv`? The file name and location. To make our lives simpler here, we'll use the URL and read it directly from my website. 

```{r load-data, message=FALSE, warning=FALSE}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")

nplayers <- nrow(players)

avgminutes <- players %>%
  group_by(Age) %>%
  summarise(
    count = n(),
    mean_minutes = mean(MP),
    median_minutes = median(MP)
  ) %>%
  arrange(desc(mean_minutes)) %>% slice(1)

threesleader <- players %>% arrange(desc(`3P`)) %>% slice(1)
```
```{r load-data-exercise, exercise = TRUE}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")
```
```{r load-data-exercise-solution}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")
```
```{r load-data-exercise-check}
grade_this_code()
```

### Exercise 1: Using head to see your data 

Now we can inspect the data we imported. What does it look like? What's in it? What do we have to work with? 

To do that, we use `head()` and **put the name of the variable we created above between the parenthesis** to show the headers and **the first six rows of data**. 

```{r head-data, exercise=TRUE, exercise.setup = "load-data"}
head(??????)
```
```{r head-data-solution, exercise.reveal_solution = FALSE}
head(players)
```
```{r head-data-check}
grade_this_code()
```
<div id="head-data-hint">
**Hint:** The thing you need is to the left of a <- in a block above.
</div>

Let's look at this. As you can see by the data, we have six players, **which is what we expect** from `head()`. But notice the first row -- the headers. We have things like `team` and `pos` for team and position. And we have a lot of number columns, showing everything from matches started (ms) to kills. 

### Exercise 2: Using glimpse to see your data 

Why do this? Why use head? It gives us a glimpse of the data. Which ... conveniently ... is also a way to look at your data, and it works the same way as `head()` 

```{r glimpse-data, exercise=TRUE, exercise.setup = "load-data"}
glimpse(??????)
```
```{r glimpse-data-solution, exercise.reveal_solution = FALSE}
glimpse(players)
```
```{r glimpse-data-check}
grade_this_code()
```
<div id="glimpse-data-hint">
**Hint:** This works the same as head() above.
</div>

Either of these will guide you with what you have to work with in the coming problems. 

## Answering questions with code

```{r numberplayers, exercise=FALSE, exercise.eval=TRUE, exercise.setup = "load-data", results='asis'}
glue("What if we wanted to know how many players there were by position, i.e. how many guards are there have been in the NBA? Or centers? To do that by hand, we'd have to take each of the {nplayers} records and sort them into piles. Forwards over here. Shooting guards over here. We'd put them in groups and then count them.")
```

`dplyr` has a **group by** function in it that does just this. A massive amount of data analysis involves grouping like things together at some point. So it's a good place to start.

**The secret to writing code is that much of it is a pattern.** The building blocks largely stay the same. With the tidyverse, this is *especially* true.

To accomplish our goal, we start with the name of the data we created before. The same name we used in `head()` and `glipse()`. We'll take that data and we'll introduce a new operator: `%>%`. The best way to read that operator, in my opinion, is to interpret that as "and then do this." So the start of our pattern is `data %>%` which you can read as `data and now we'll do this`.

The this, in our case, is `group_by()`. Like `head()` and `glimpse()`, **we need to put something in the parenthesis**. But unlike `head()` and `glimpse()`, this time the something comes from that first row of head. We are grouping data by one of the pieces of the data -- a field. *If we're trying to group by position*, which field looks like it holds the position name? Let's use head again and take a look at the very top row in bold.

```{r head-review, exercise=TRUE, exercise.setup = "load-data"}
head(??????)
```
```{r head-review-solution, exercise.reveal_solution = FALSE}
head(players)
```
```{r head-review-check}
grade_this_code()
```

That block of code you just ran has two hints for the code block you'll have to complete later: What data are you using? What field in that data do you want to group by? Keep those in mind.

### Exercise 3: Group by and count

After we group our data together by the thing we want to group it by, we need to count how many things are in each group. We do that first by saying we want to summarize our data (a count is a part of a summary). To get a summary, we have to tell it what we want to summarize. So in this case, we want a count. To get that, let's create a thing called total and set it equal to n(), which is `dplyr`s way of counting something. 

Here's the pattern. **You fill in where there are all capitals with what needs to be there.** What you fill in are the two hints from above.

```{r group-by, exercise=TRUE, exercise.setup = "load-data", message=FALSE}
NAMEOFDATA %>%
  group_by(NAMEOFFIELD) %>%
  summarise(
    total = n()
  ) 
```
```{r group-by-solution, exercise.reveal_solution = FALSE}
players %>%
  group_by(Pos) %>%
  summarise(
    total = n()
  ) 
```
```{r group-by-check}
grade_this_code()
```
<div id="group-by-hint">
**Hint:** The name of your data is the same as what you used in `head()` and `glimpse()`. Then, look carefully at the first row of the output of head. What you need is in bold at the top.
</div>
In this case, we wanted to group together positions, signified by the field name `Pos`. After we group the data, we need to count them up. In dplyr, we use `summarize` [which can do more than just count things](http://dplyr.tidyverse.org/reference/summarise.html). Inside the parentheses in summarize, we set up the summaries we want. In this case, we just want a count of the positions: `total = n(),` says create a new field, called `total` and set it equal to `n()`, which might look weird, but it's common in stats. The number of things in a dataset? Statisticians call in n. There are n number of players in this dataset. So `n()` is a function that counts the number of things there are. 

### Exercise 4: Arranging data

And when we run that, we get a list of positions with a count next to them. But it's not in any order. So we'll add another And Then Do This %>% and use `arrange`. Arrange does what you think it does -- it arranges data in order. By default, it's in ascending order -- smallest to largest. But if we want to know the position with the most players, we need to sort it in descending order. In arrange, we need to tell it what we are arranging by. **If you have group_by, then whatever you are arranging by will *always* come from inside summarize.** The pattern looks like this:

```{r group-by-2, exercise=TRUE, exercise.setup = "load-data", message=FALSE}
players %>%
  group_by(???) %>%
  summarise(
    total = n()
  ) %>% 
  arrange(desc(?????))
```
```{r group-by-2-solution, exercise.reveal_solution = FALSE}
players %>%
  group_by(Pos) %>%
  summarise(
    total = n()
  ) %>% 
  arrange(desc(total))
```
```{r group-by-2-check}
grade_this_code()
```
<div id="group-by-2-hint">
**Hint:** The only thing new here is the arrange, and what goes in arrange comes from summarize, and it isn't the n() parts.
</div>
What is the most common position in volleyball? If you look at the row with the largest total, you get your answer. 

### Exercise 5: Grouping by more than one thing 

We can, if we want, group by more than one thing. Which team has the most of a single position in a single season? To do that, we can group by the team, season and position in the data:

```{r group-by-3, exercise=TRUE, exercise.setup = "load-data", message=FALSE}
players %>%
  group_by(????, ??????, ???) %>%
  summarise(
    total = n()
  ) %>% 
  arrange(desc(?????))
```
```{r group-by-3-solution, exercise.reveal_solution = FALSE}
players %>%
  group_by(Tm, Season, Pos) %>%
  summarise(
    total = n()
  ) %>% 
  arrange(desc(total))
```
```{r group-by-3-check}
grade_this_code()
```
<div id="group-by-3-hint">
**Hint:** If you're reading this, you're going too fast. I told you what to put in group_by in the sentence above this code window. I will do that a lot this semester. Get used to reading carefully. Hints are EVERYWHERE. 
</div>

You can learn a bit about how teams are assembled by looking at these simple counts.

### Exercise 6: More in summarize 

In the last example, we grouped some data together and counted it up, but there's so much more you can do. **You can do multiple measures in a single step as well.**

Sticking with our NBA player data, we can calculate any number of measures inside summarize. Here, we'll use R's built in `mean` and `median` functions to calculate ... well, you get the idea.

Let's look just a the number of minutes (MP) each age gets. **In summarize, what you do left of the equal sign is giving something a name**. We should name our new columns what they are: `mean_minutes` and `median_minutes`. Then let's arrange it by the average number of minutes (hint: that wording, average number of minutes, is a trick. You arrange by the name of a column created in summarize). 

```{r group-by-4, exercise=TRUE, exercise.setup = "load-data"}
???????? %>%
  group_by(???) %>%
  summarise(
    count = n(),
    ????????? = mean(?),
    ??????????? = median(?)
  ) %>%
  arrange(desc(?????????))
```
```{r group-by-4-solution, exercise.reveal_solution = FALSE}
players %>%
  group_by(Age) %>%
  summarise(
    count = n(),
    mean_minutes = mean(MP),
    median_minutes = median(MP)
  ) %>%
  arrange(desc(mean_minutes))
```
```{r group-by-4-check}
grade_this_code()
```
<div id="group-by-4-hint">
**Hint:** What goes in `group_by`, `mean()` and `median()` all come from column names you find with `head()`. What goes before the = are names in the previous paragraph.
</div>

```{r avgminutes, exercise=FALSE, exercise.eval=TRUE, exercise.setup = "load-data", results='asis'}
glue("So there's {nplayers} players in the data. Of them, {avgminutes$count} are {avgminutes$Age} years old, which if you know professional basketball, that kinda fits. Players peak in their late 20s. The average {avgminutes$Age}-year old plays in {format(avgminutes$mean_minutes, digits=1)} minutes on average. The median -- where half are above and half are below -- is {avgminutes$median_minutes}.")
```

### Exercise 7: Arranging two ways

One thing to keep in mind is this: You can use any of these verbs -- group_by, summarize, arrange -- on their own without the others. For example, what player had the most three point makes since Bird and Magic entered the scene? To find this out, we don't need a group_by or summarize, so what goes in arrange is just the field name. Let's put this in order of 3P, which we'll need to wrap in backticks because column names that start with a number are a no-no. 

First let's get the most kills.

```{r group-by-5, exercise=TRUE, exercise.setup = "load-data"}
players %>% arrange(desc(`??`))
```
```{r group-by-5-solution, exercise.reveal_solution = FALSE}
players %>% arrange(desc(`3P`))
```
```{r group-by-5-check}
grade_this_code()
```

Now we can get the least by removing desc() and just doing arrange by our field name. 

```{r group-by-6, exercise=TRUE, exercise.setup = "load-data"}
players %>% arrange(`3P`)
```
```{r group-by-6-solution, exercise.reveal_solution = FALSE}
players %>% arrange(`3P`)
```
```{r group-by-6-check}
grade_this_code()
```

```{r killer, exercise=FALSE, exercise.eval=TRUE, exercise.setup = "load-data", results='asis'}
glue("The single season three-ball leader since 1979-80? {threesleader$Player} of the {threesleader$Tm} with {threesleader$`3P`} kills. The lowest? A whole lot of people with 0.")
```

That's a huge difference.

So when choosing a measure of the middle, you have to ask yourself -- could I have extremes? Because a median won't be sensitive to extremes. It will be the point at which half the numbers are above and half are below. The average or mean will be a measure of the middle, but if you have a bunch of pine riders and then one iron superstar, the average could be wildly skewed.

We'll work more on that in the future.

### Exercise 8: Even more aggregates

There's a ton of things we can do in summarize -- we'll work with more of them as the course progresses -- but here's a few other questions you can ask.

Which position in the NBA generates the most points? And what is the highest and lowest point total for that position? And how wide is the spread between points? We can find that with `sum` to add up the points to get the total points, `min` to find the minimum points, `max` to find the maximum points and `sd` to find the standard deviation in the numbers. 

```{r group-by-7, exercise=TRUE, exercise.setup = "load-data"}
players %>% 
  group_by(???) %>% 
  summarise(
    total = sum(PTS), 
    avg_points = mean(PTS), 
    min_points = min(PTS),
    max_points = max(PTS),
    stdev_points = sd(PTS)) %>% 
  arrange(desc(total))
```
```{r group-by-7-solution, exercise.reveal_solution = FALSE}
players %>% 
  group_by(Pos) %>% 
  summarise(
    total = sum(PTS), 
    avg_points = mean(PTS), 
    min_points = min(PTS),
    max_points = max(PTS),
    stdev_points = sd(PTS)) %>% 
  arrange(desc(total))
```
```{r group-by-7-check}
grade_this_code()
```
<div id="group-by-7-hint">
**Hint:** Breathe deep. Slow down. Think about it. What's missing here is stuff you've done already. 
</div>

Most of the stats here we can intuitively understand -- a minimum of zero means they probably didn't play at all. A large max means that player was good at basketball. The one that you should give an extra check is the standard deviation. When your standard deviation is large, that's a sign you've got some huge points players and a bunch of bench players and you should consider which measure of the middle you're going to use. 

In data analysis, `group_by` and `summarize` are two of the most basic, but most common functions. With these functions, you can take every game played and turn it into season totals, team statistics and many others. You can take every pitch thrown and look at them by type. What you need to do all this you learned here. What remains is creative application of what you've learned.