---
title: "Advanced Sports Data Lesson 3: Filter review"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to narrow in on what's important and remove what isn't.
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(tidyverse)
library(glue)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.completion=FALSE)
```

# Filtering

## The basics

More often than not, we have more data than we want. Sometimes we need to be rid of that data. In `dplyr`, there's two ways to go about this: filtering and selecting.

**Filtering creates a subset of the data based on criteria**. All records where the count is greater than 10. All records that match "Nebraska". Something like that. 

**Selecting simply returns only the fields named**. So if you only want to see School and Attendance, you select those fields. When you look at your data again, you'll have two columns. If you try to use one of your columns that you had before you used select, you'll get an error.  

Let's continue our work on NBA facilitators from last season. You can download the data here if you want to use it in your own notebook -- and this is the same data we used in the mutating tutorial, so if you already downloaded it, you don't need to do it again. **For purposes of this exercise, you don't need to do this. The data is included here if you want to try this in your own notebook.**

```{r echo=FALSE, class.output="bg-info", results="asis",  message=FALSE,  warning=FALSE}
library(downloadthis)
library(glue)

dllink <- download_link(
  link = "http://mattwaite.github.io/sportsdatafiles/allplayers.csv",
  button_label = "Download csv file",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)

glue("<pre><p><strong>For this walkthrough:</strong></p><p>{dllink}</p></pre>")

```

First we'll need the tidyverse. Your first step is always loading libraries and you'll need to run this step in nearly every single thing you do.

```{r load-tidyverse, exercise=TRUE}
library(tidyverse)
```
```{r load-tidyverse-solution}
library(tidyverse)
```
```{r load-tidyverse-check}
grade_this_code()
```

Now import the data.

```{r filtering-load-data, message=FALSE, warning=FALSE}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")

topcleanpass <- players %>%
  filter(MP >= 2600) %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover)) %>% slice(1)

cleanpass <- players %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover))
```
```{r filtering-load-data-exercise, exercise = TRUE}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")
```
```{r filtering-load-data-exercise-solution}
players <- read_csv("http://mattwaite.github.io/sportsdatafiles/allplayers.csv")
```
```{r filtering-load-data-exercise-check}
grade_this_code()
```

So, first things first, let's say we don't care about old players. You're children of the 2000s, you don't care about the 80s and 90s. 

Before we do anything, a note about conditions. Most of the conditional operators you'll understand -- greater than and less than are > and <. You can add an equal sign to > and < to get greater than OR equal to (or the opposite). So >= would give you all whatevers greater than or equal to that number. 

The tough one to remember is equal to. In conditional statements, equal to is == not =. If you haven't noticed, = is a variable assignment operator, not a conditional statement. So equal is == and NOT equal is !=. 

So if you want to see games from the 2000s and beyond, what would you do?

First, let's use head to give us a peek at the data. This data's layout is ... not great ... so you're going to want to scroll all the way to the right to see the most useful columns for this exercise.

```{r head-data, exercise=TRUE, exercise.setup = "filtering-load-data"}
head(?????)
```
```{r head-data-solution}
head(players)
```
```{r head-data-check}
grade_this_code()
```

Now that we can see the column names, we now know Season is where the years are. Replace the all caps and question marks here and see if you can get only 2000s players.

### Exercise 1: Current players only please 

```{r filtering-new, exercise=TRUE, exercise.setup = "filtering-load-data"}
NAMEOFDATA %>% filter(NAMEOFCOLUMN ?? 2000)
```
```{r filtering-new-solution, exercise.reveal_solution = FALSE}
players %>% filter(Season >= 2000)
```
```{r filtering-new-check}
grade_this_code()
```

And what do you see? Lots of dudes who have played more recently than Magic Johnson.

### Exercise 2: So who really is delivering the ball?

Now we get back to what we started last time. When last we left off, we were looking at the assist/turnover ratio and the answer we came up with was ... not someone you've ever heard of. Not one of the best passers in the game, that's for sure. 

So who is? This is an important concept to get. With some metrics, there needs to be a cutoff. There needs be some threshold to be included. What happens to our list if we only include people who played 2600 minutes or more? 

```{r filtering-passers, exercise=TRUE, exercise.setup = "filtering-load-data"}
players %>%
  filter(?? ?? 2600) %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover))
```
```{r filtering-passers-solution, exercise.reveal_solution = FALSE}
players %>%
  filter(MP >= 2600) %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover))
```
```{r filtering-passers-check}
grade_this_code()
```

```{r passers, exercise=FALSE, exercise.eval=TRUE, exercise.setup = "filtering-load-data", results='asis'}
glue("It's impossible to explain {topcleanpass$Player}. You can see his name a bunch in the first 10 rows of data. But what you don't see is he was 5'3\" tall -- literally the shortest player to every play in the NBA. And he was GOOD.")
```

We can see, though, that all of these players are point guards, which you would expect. What if we looked at guards who were supposed to be scoring? Aka shooting guards. There's a filter for that too. 

Note: The position name *must* be exact. Sg or ShootingGuard. You have to spell it and capitalize it *exactly* like it is in the data. 

### Exercise 3: Finding shooting guards

```{r filtering-big, exercise=TRUE, exercise.setup = "filtering-load-data"}
players %>%
  filter(?? ?? 2600) %>%
  filter(??? == "??") %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover))
```
```{r filtering-big-solution, exercise.reveal_solution = FALSE}
players %>%
  filter(MP >= 2600) %>%
  filter(Pos == "SG") %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover))
```
```{r filtering-big-check}
grade_this_code()
```

Aside from the top name, look at the 9th person on the list. Was Magic a shooting guard or did he run the point? If you know Magic's career ... it's a complicated question. Something to look out for in the future is how players who do a lot get classified.

## Selecting data to make it easier to read

With all of this, we have a lot of extra fields or columns we don't want. When we're talking about facilitators, maybe we only want to see the player name, their position, their age, team, games, assists, turnovers and the assist turnover ratio. We can do that with select, and with this class, you're going to be using select a bunch, because the algorithms we're going to use just want us to focus on certain things. 

### Exercise 4: Select to simplify 

Generally speaking, select is something we want to place carefully. We don't want to select first, and then need someting. Often, since select is mainly cosmetic, we would do it last.

```{r selecting-power, exercise=TRUE, exercise.setup = "filtering-load-data"}
players %>%
  filter(?? >= 2600) %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover)) %>%
  select(
    Player,
    Pos,
    Age,
    Tm,
    ???,
    ???,
    ??????????????
  )
```
```{r selecting-power-solution, exercise.reveal_solution = FALSE}
players %>%
  filter(MP >= 2600) %>%
  mutate(AssistTurnover = (AST/TOV)
) %>%
  arrange(desc(AssistTurnover)) %>%
  select(
    Player,
    Pos,
    Age,
    Tm,
    AST,
    TOV,
    AssistTurnover
  )
```
```{r selecting-power-check}
grade_this_code()
```